Better Splitting Algorithms for Parallel Corpus Processing

### 2012-08-06 doc/latex/paper.tex
Corrected naive Algorithm 1, per request by Reviewer 2.
Camera-ready changes

@@

Values for $\bm{\lambda}$ are obtained by optimizing an objective function such as BLEU \citep{papinenietal01} against a development set, most commonly using minimum error rate training (MERT) \citep{och03}. Each iteration of MERT requires this development set to be re-translated using a new set of $\bm{\lambda}$ weights. MERT is one of the slowest components in a typical machine translation training pipeline, and translating the development set is nearly always the slowest step in MERT. We now examine techniques for speeding up MERT by distributing translation jobs across a cluster of parallel computational nodes.

Ideally, all parts should take the same amount of time to translate. While naive splitting techniques reduce the time required for each translation iteration by splitting the work between $p$ computational nodes, in practice some parts may take much longer to complete than others. This can lead to significant computational slack time.
%
To address this problem, we develop three novel algorithms for splitting translation tasks in a parallel computing environment, drawing on research in parallel machine scheduling.

@@

While the models in Equation \ref{eq:smt} could, in theory, condition on previously translated sentences, in practice virtually no widely used models do so. It is therefore very straightforward to split the data into $p$ parts, and translate each part independently on $p$ computational nodes. Scripts implemented in Moses \citep{moses} do exactly that, simply splitting the data into $p$ arbitrary parts such that each part contains the same number of lines (Algorithm \ref{alg:naive}).

\begin{algorithm}
\caption{Split input text into $n$ parts such that each part contains the same number of lines.}
\begin{algorithmic}
\fontspec{TeX Gyre Pagella}
\Function{\fontspec{TeX Gyre Pagella}\textsc{Naive-Split}}{$n$,input}
\State $\ell \gets n / $ input.length
\For {$p \gets 0 \ldots (n-1)$}
\State $i \gets \ell \times p$
\For {$j \gets i \ldots (i+\ell-1)$}
\If{$j < \ell$}
\State output[$p$].append(input[$j$])
\Else
\State \textbf{break}
\EndIf
\EndFor
\EndFor
\State \textbf{return} output
\EndFunction
\end{algorithmic}
\label{alg:naive}
\end{algorithm}

@@

@@

\begin{figure*}[p]
\centering
\subfloat[Decoding times in \textbf{seconds} for decoder configured using a 5-gram language model.]{
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={160,150,140,130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Seconds),
ymin=0,
%enlarge x limits=0.5,
%scale only axis=false,
%axis equal=false,
%legend style={
%	at={(0.5,-0.15)},
%	anchor=north,legend columns=-1
%},
%ybar=0pt,
width=13cm,
height=7cm,
ybar,
bar width=7pt,
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
%bar shift=3pt
%ybar interval=0.7
]
%
\input{ngram/ngram.naive-Maximum_times}
\input{ngram/ngram.histogram-Maximum_times}
\input{ngram/ngram.words-Maximum_times}
\input{ngram/ngram.time-Maximum_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%\caption{\fontspec{TeX Gyre Pagella}Decoding times (in {\textbf{seconds}}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a 5-gram language model.}
%\label{fig:maxspeed}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%

%\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[p]
%\centering
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}%[scale=2]
\begin{axis}[
xmode=log,
ymode=normal,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Hours),
ymin=0,
%ybar=0pt,
ybar,
bar width=7pt,
%bar shift=0pt,
width=13cm,
height=7cm,
%cycle list name=black white
%colormap name=bluered
%cycle list={fill={blue!80!red!30}, fill=red!66, fill=yellow!33, black}
%cycle list={fill=blue!30!white,fill=red!60!white,fill=brown!90!white,fill=black}
%cycle list={fill=blue!30!white,fill=red!30!white,fill=brown!30!white,fill=grey}
%cycle list={fill=gray!25,fill=gray!50,fill=gray!75,fill=gray!100}
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
%cycle list name=color
]
%
\input{hhmm/hhmm.naive-Maximum_times}
\input{hhmm/hhmm.histogram-Maximum_times}
\input{hhmm/hhmm.words-Maximum_times}
\input{hhmm/hhmm.time-Maximum_times}
%
%\legend{Naive,Histogram,Words,Oracle}
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%\caption{\fontspec{TeX Gyre Pagella}Decoding times (in \textbf{hours}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.}
\caption{\fontspec{TeX Gyre Pagella}Decoding times for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}).}
\label{fig:maxspeed}
\end{figure*}

@@

@@

Implementations of all four algorithms in Ruby and Java are provided at \url{http://github.com/dowobeha/balance-corpus}.

@@

\clearpage
\thanksnostar{Opinions, interpretations, conclusions, and recommendations are those of the authors and are not necessarily endorsed by the United States Air Force. Cleared for public release on 19 Jan 2011. Originator reference number RH-11-107802. Case number 88ABW-2011-0185.}

>>>>

\defaultfontfeatures{Mapping=tex-text}

@@

Values for $\bm{\lambda}$ are obtained by optimizing an objective function such as BLEU \citep{papinenietal01} against a development set, most commonly using minimum error rate training (MERT) \citep{och03}. Each iteration of MERT requires this development set to be re-translated using a new set of $\bm{\lambda}$ weights. MERT is one of the slowest components in a typical machine translation training pipeline, and translating the development set is nearly always the slowest step in MERT. We now examine techniques for speeding up the translation process by splitting a source document into parts and distributing the translation of those parts across parallel computational nodes.

Ideally, all parts should take the same amount of time to translate. While naive splitting techniques reduce the time required for each translation iteration by splitting the work between $n$ computational nodes, in practice some parts may take much longer to complete than others. This can lead to significant computational slack time.
%
To address this problem, we develop three novel algorithms for splitting translation tasks in a parallel computing environment, drawing on research in parallel machine scheduling.

@@

@@

We now briefly examine the existing research most relevant to our task. \citet{Hu61} and Graham \citeyearpar{Graham66,Graham69} develop various list scheduling algorithms. This family of algorithms prioritizes jobs into a queue, then assigns jobs to machines in queue order. This approach attempts to evenly balance the load on each execution host \citep{DeMorton80,ChengSin90}. 
%Of the techniques which we examine in this work, Algorithms \ref{alg:naive}, \ref{alg:words}, and \ref{alg:times} fall into this category.
Both Algorithm \ref{alg:naive} below and the techniques we develop in Section \ref{sec:better} fall into this family of algorithms.

\begin{algorithm}
\caption{Split input text into $n$ parts such that each part contains the same number of lines. In cases where the total number of lines is not evenly divisible by $n$, the last part will contain fewer lines than each of the other parts.}
\begin{algorithmic}
\fontspec{TeX Gyre Pagella}
\Function{\fontspec{TeX Gyre Pagella}\textsc{Naive-Split}}{$n$,input}
\State $s \gets $ input.length
\State $\ell \gets \lceil s / n\rceil$
\For {$p \gets 0 \ldots (n-1)$}
\State $i \gets \ell \times p$
\For {$j \gets i \ldots \min(i+\ell-1$,$s-1)$}
%\If{$j < \ell$}
\State output[$p$].append(input[$j$])
%\Else
%\State \textbf{break}
%\EndIf
\EndFor
\EndFor
\State \textbf{return} output
\EndFunction
\end{algorithmic}
\label{alg:naive}
\end{algorithm}

While the models in Equation \ref{eq:smt} could, in theory, condition on previously translated sentences, in practice virtually no widely used models do so. It is therefore very straightforward to split the data into $n$ parts, and translate each part independently on $n$ computational nodes.

@@

\begin{figure*}
\centering
\begin{tikzpicture}
\begin{axis}[
xmode=normal,
ymode=normal,
xtick={100,200,300,400,500,600,700,800},
ytick={0,50,100,150,200,250,300},
tick label style={font=\tiny},
xlabel=Sentence Number,
ylabel=Sentence Length,
ymin=0,
xmin=0,
xmax=883,
xtick pos=left,
ytick pos=left,
xtick align=outside,
ytick align=outside,
ybar,
bar width=0.5pt,
cycle list={fill=black},
width=13cm,
height=5cm
]
%\input{sentence_lengths}
\input{figures/sentence_lengths/Corpus.nist08_dev/plot.tex}
\end{axis}
\end{tikzpicture}
\caption{Sentence length (in words) of sentences from the NIST OpenMT 2008 Urdu-English task.}
\label{fig:sentence-length}
\end{figure*}

% \begin{figure*}[!h]
% \centering
% \begin{tikzpicture}
% \begin{axis}[
% xmode=normal,
% ymode=normal,
% xtick={100,200,300,400,500,600,700,800},
% ytick={0,5,10,15},
% tick label style={font=\tiny},
% %xlabel=Sentence Number,
% ylabel=Translation Time,
% ymin=0,
% xmin=0,
% xmax=883,
% xtick pos=left,
% ytick pos=left,
% xtick align=outside,
% ytick align=outside,
% ybar,
% bar width=0.5pt,
% cycle list={fill=black},
% width=13cm,
% height=5cm
% ]
% \input{sentence_times}
% \end{axis}
% \end{tikzpicture}
% \caption{Translation decoding time (in seconds) of each sentence in the development set of the NIST OpenMT 2008 Urdu-English task for a decoder configured using a 5-gram language model.}
% \label{fig:sentence-times}
% \end{figure*}

\begin{figure*}[!p]
\centering
\subfloat[Algorithm \ref{alg:naive} --- \textsc{Naive-Split}]{
\begin{tikzpicture}
\begin{axis}[
xmode=normal,
ymode=normal,
xtick={0,8,16,24,32,40,48,56,64},%data,%{100,200,300,400,500,600,700,800},
ytick={0,10,20,30,40,50,60,70,80},
tick label style={font=\tiny},
%xlabel=Split Part Number,
ylabel=Translation Time,
ymin=0,
ymax=80,
xmin=0,
xmax=64,
xtick pos=left,
ytick pos=left,
xtick align=outside,
ytick align=outside,
ybar,
bar width=5pt,
cycle list={fill=black},
width=13cm,
height=4cm
]
%\input{split/time_sum.64}
\input{figures/parts_time_summary/Algorithm.naive+Corpus.nist08_dev+Parts.64/plot.tex}
\end{axis}
\end{tikzpicture}
%\caption{Total time (in seconds) for each part when split using Algorithm \ref{alg:naive}.}
%\caption{Translation decoding time (in seconds) of sentences from the NIST OpenMT 2008 Urdu-English task with a decoder configured using a 5-gram language model.}
\label{fig:plot-naive}
%\end{figure*}
}

\subfloat[Algorithm \ref{alg:histogram} --- \textsc{Histogram-Split}]{
%\begin{figure*}[!p]
%\centering
\begin{tikzpicture}
\begin{axis}[
xmode=normal,
ymode=normal,
xtick={0,8,16,24,32,40,48,56,64},%data,%{100,200,300,400,500,600,700,800},
ytick={0,10,20,30,40,50,60,70,80},
tick label style={font=\tiny},
%xlabel=Split Part Number,
ylabel=Translation Time,
ymin=0,
ymax=80,
xmin=0,
xmax=64,
xtick pos=left,
ytick pos=left,
xtick align=outside,
ytick align=outside,
ybar,
bar width=5pt,
cycle list={fill=black},
width=13cm,
height=4cm
]
%\input{split/histogram_time_sum.64}
\input{figures/parts_time_summary/Algorithm.histogram+Corpus.nist08_dev+Parts.64/plot.tex}
\end{axis}
\end{tikzpicture}
\label{fig:plot-histogram}
%\caption{Total time (in seconds) for each part when split using Algorithm \ref{alg:histogram}.}
%\caption{Translation decoding time (in seconds) of sentences from the NIST OpenMT 2008 Urdu-English task with a decoder configured using a 5-gram language model.}
%\label{fig:histogram-split-times}
}
%\end{figure*}

\subfloat[Algorithm \ref{alg:words} --- \textsc{Words-Split}]{
%\begin{figure*}[!p]
%\centering
\begin{tikzpicture}
\begin{axis}[
xmode=normal,
ymode=normal,
xtick={0,8,16,24,32,40,48,56,64},%data,%{100,200,300,400,500,600,700,800},
ytick={0,10,20,30,40,50,60,70,80},
tick label style={font=\tiny},
%xlabel=Split Part Number,
ylabel=Translation Time,
ymin=0,
ymax=80,
xmin=0,
xmax=64,
xtick pos=left,
ytick pos=left,
xtick align=outside,
ytick align=outside,
ybar,
bar width=5pt,
cycle list={fill=black},
width=13cm,
height=4cm
]
%\input{split/histogram_time_sum.64}
\input{figures/parts_time_summary/Algorithm.words+Corpus.nist08_dev+Parts.64/plot.tex}
\end{axis}
\end{tikzpicture}
\label{fig:plot-words}
%\caption{Total time (in seconds) for each part when split using Algorithm \ref{alg:words}.}
%\caption{Translation decoding time (in seconds) of sentences from the NIST OpenMT 2008 Urdu-English task with a decoder configured using a 5-gram language model.}
%\label{fig:words-split-times}
%\end{figure*}
}

\subfloat[Algorithm \ref{alg:times} --- \textsc{Times-Split}]{
%\begin{figure*}[!p]
%\centering
\begin{tikzpicture}
\begin{axis}[
xmode=normal,
ymode=normal,
xtick={0,8,16,24,32,40,48,56,64},%data,%{100,200,300,400,500,600,700,800},
ytick={0,10,20,30,40,50,60,70,80},
tick label style={font=\tiny},
%xlabel=Split Part Number,
ylabel=Translation Time,
ymin=0,
ymax=80,
xmin=0,
xmax=64,
xtick pos=left,
ytick pos=left,
xtick align=outside,
ytick align=outside,
ybar,
bar width=5pt,
cycle list={fill=black},
width=13cm,
height=4cm
]
%\input{split/histogram_time_sum.64}
\input{figures/parts_time_summary/Algorithm.time+Corpus.nist08_dev+Parts.64/plot.tex}
\end{axis}
\end{tikzpicture}
\label{fig:plot-time}
}
\caption{Total translation time (in \textbf{seconds}) for each part when data from the NIST OpenMT 2008 Urdu-English task is split into 64 parts using each of four algorithms.}
%\caption{Translation decoding time (in seconds) of sentences from the NIST OpenMT 2008 Urdu-English task with a decoder configured using a 5-gram language model.}
\label{fig:time-split-times}
\end{figure*}

\section{Better Splitting Algorithms}
\label{sec:better}

We begin by examining the development set for the NIST OpenMT 2008 Urdu-English task. When tuning the model weights $\bm{\lambda}$, this development set will be translated numerous times. We observe in Figure \ref{fig:sentence-length} that the number of words in each sentence varies widely and unevenly throughout the corpus. %This is notable in that when this corpus is split using Algorithm \ref{alg:naive},

%Scripts included with Moses \citep{moses} do exactly that, simply splitting the data into $p$ arbitrary parts (Algorithm \ref{alg:naive}) of $\ell$ or fewer lines each. 
Using Algorithm \ref{alg:naive}, scripts included with Moses \citep{moses} split the corpus into $n$ parts of of $\ell$ or fewer lines each; $\ell$ is the smallest integer greater than or equal to $s/n$, where $s$ is the total number of input sentences. The first $\ell$ lines comprise the first part, the next $\ell$ lines comprise the second part, and so on. Thus, each part contains exactly $\ell$ lines, with the possible exception of the last part, which contains fewer than $\ell$ lines when $s$ is not evenly divisible by $n$.

Figure \ref{fig:plot-naive} shows the amount of time taken to translate each part of the development set, split using Algorithm \ref{alg:naive} into 64 parts, using Moses configured with a 5-gram language model. 
We observe that the shape of Figure \ref{fig:plot-naive} generally matches that of Figure \ref{fig:sentence-length}.
%Because Algorithm \ref{alg:naive} partitions contiguous sentences together, and because phrase-based decoding with a finite distortion limit takes linear time, it is not surprising to see that the shape of the graph in Figure \ref{fig:plot-naive} generally matches that of \ref{fig:sentence-length}. 
There is substantial variance in translation time between the parts in Figure \ref{fig:plot-naive}, with some parts taking nearly 80 seconds and many others finishing in well under 10 seconds. We observe that this disparity is largely due to an imbalance of short versus long sentences between the parts. Because short sentences take less time to translate than long sentences, parts assigned mostly short sentences finish much faster than parts that are assigned many longer sentences.

%Algorithm \ref{alg:naive} first calculates

%In examining the logs of various translations jobs split using Algorithm \ref{alg:naive}, we observe that slack time results primarily from situations where some jobs are assigned a disproportionate number of short sentences. 
%

To remedy this imbalance, we propose Algorithm \ref{alg:histogram}. 
%
Prior to splitting the data into parts, Algorithm \ref{alg:histogram} begins by examining the number of words in each sentence.
%
Sentences are sorted according to length, then assigned in turns to parts. 
%
This round-robin distribution of sentences into parts results in the sentence length histograms for each part being approximately equal. 
%
While Algorithm \ref{alg:histogram} attempts to balance short and long sentences across parts, we nevertheless observe non-trivial imbalance in translation times across parts in Figure \ref{fig:plot-histogram}.

% \begin{figure*}
% \centering
% \begin{tikzpicture}
% \begin{axis}[
% xmode=normal,
% ymode=normal,
% xtick={10,20,30,40,50,100,150,200,250,300},
% ytick={1,5,10,15,20,25,30,35},
% tick label style={font=\tiny},
% xlabel=Sentence Length (in words),
% ylabel=Number of Sentences,
% ymin=0,
% xmin=0,
% xtick pos=left,
% ytick pos=left,
% xtick align=outside,
% ytick align=outside,
% ybar,
% bar width=0.5pt,
% cycle list={fill=black},
% width=13cm,
% height=7cm
% ]
% \input{devset.histogram.part1}
% \end{axis}
% \end{tikzpicture}
% \caption{Sentence length histogram for the development set of the NIST OpenMT 2008 Urdu-English task.}
% \label{fig:histogram}
% \end{figure*}

%While Algorithm \ref{alg:histogram} attempts to balance short and long sentences across parts, this algorithm can still lead to sizeable slack time in cases with severe length imbalances between sentences. Development set for the NIST OpenMT 2008 Urdu-English task 
%
%we may be able to achieve even more balanced runtimes by balancing the total number of words in each part.

\begin{algorithm}
\caption{Split input text into $n$ parts to balance the number of words for all parts.}
\begin{algorithmic}
\fontspec{TeX Gyre Pagella}
\Function{\fontspec{TeX Gyre Pagella}\textsc{Words-Split}}{$n$,input}
\For {$i \gets 0 \ldots ($ input.length $-1)$}
\State sentence[$i$].length $\gets$ input[$i$].length
\State sentence[$i$].index $\gets i$
\EndFor
\State \Call{Sort}{sentence} $\left\{|x,y|\ y.\mbox{length}\Leftrightarrow x.\mbox{length}\right\}$ 
\State \Comment{Sort sentences by length, in reverse order}
\For {$i \gets 0 \ldots ($ input.length $-1)$}
\State $p \gets$ \Call{Least}{words}
\State \Comment{Find partition with fewest words}
\State output[$p$].append(input[sentence[$i$].index])
\State words[$p$] $\gets$\hspace{0em}words[$p$] $+$ sentence[$i$].length%
%\State words[$p$]\hspace{0em}$\gets$\hspace{0em}words[$p$]\hspace{0em}$+$\hspace{0em}sentence[$i$].length%
\EndFor
\State \textbf{return} output
\EndFunction
\end{algorithmic}
\label{alg:words}
\end{algorithm}
%
%

%
%
To improve this remaining imbalance, we propose Algorithm \ref{alg:words}.
%
In Algorithm \ref{alg:words}, sentences are sorted by length into a queue, with longest sentences at the head of the queue.
%Algorithm \ref{alg:words} counts the number of words in each sentence to be translated. Sentences are then ordered by this length, in reverse order. Initially, no computation slots have assigned translation jobs. 
%
Initially, no sentences have been assigned to any part.
%
The longest sentence, at the head of the queue, is assigned first to a part. As each sentence is assigned to a part, the total number of words assigned to that part is recorded. 
%
Each subsequent sentence is removed from the queue and assigned to the part with 
%
%
the least work assigned to it, as measured by number 
%
of words.
%
In Figure \ref{fig:plot-words}, we observe that most of the imbalance in translation times across parts has been resolved.

When assigning sentences to jobs, we would ideally like to know how long each sentence will take to process. Algorithms \ref{alg:histogram} and \ref{alg:words} use the number of words in each sentence as a proxy for processing time. During MERT, the same set of development sentences are translated multiple times. Since each decoding process differs only by the $\bm{\lambda}$ weights used, it is reasonable to expect little variation in decoding runtime for any given sentence across all MERT runs. With this in mind, we record the time required to translate each sentence during the first iteration of MERT. In subsequent iterations, Algorithm \ref{alg:times} uses the time recorded to translate a sentence as an estimate of the time it will take to translate that sentence again. Algorithm \ref{alg:times} differs from Algorithm \ref{alg:words} by sorting using these times instead of sentence length.
%
We see in Figure \ref{fig:plot-time} that the time imbalance between parts is virtually non-existent, with all times now within 0.01 seconds of each other.

%\todo[inline]{Describe Algorithm \ref{alg:times}. \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ }

\begin{algorithm}
\caption{Split input text into $n$ parts to balance the estimated translation time of all parts.}
\begin{algorithmic}
\fontspec{TeX Gyre Pagella}
\Function{\fontspec{TeX Gyre Pagella}\textsc{Times-Split}}{$n$,input,estimate}
\For {$i \gets 0 \ldots ($ input.length $-1)$}
\State sentence[$i$].time $\gets$ estimate[$i$]
\State sentence[$i$].index $\gets i$
\EndFor
\State \Call{Sort}{sentence} $\left\{|x,y|\ y.\mbox{time}\Leftrightarrow x.\mbox{time}\right\}$ 
\State \Comment{Sort sentences by time, in reverse order}
\For {$i \gets 0 \ldots ($ input.length $-1)$}
\State $p \gets$ \Call{Least}{times}
\State \Comment{Find partition with least time}
\State output[$p$].append(input[sentence[$i$].index])
\State times[$p$] $\gets$ times[$p$] $ + $ sentence[$i$].time
\EndFor
\State \textbf{return} output
\EndFunction
\end{algorithmic}
\label{alg:times}
\end{algorithm}

\section{Experimental Configuration}

To observe the effects of splitting algorithms on decoding speed, we translated the NIST OpenMT 2008 development set of Urdu-English data using Moses in a parallel computing cluster, distributing work using the Sun Grid Engine. We ran two decoding setups: a standard configuration using a 5-gram language model, and a much slower syntactic LM configuration following \citet{schwartzetal11}.

Figure \ref{fig:time-split-times} shows the per-part translation times for all parts of the development set when $n=64$. 
%
In Figure \ref{fig:maxspeed}, we examine the per-part translation times for only the slowest of $n$ translation jobs in each configuration for various values of $n$, ranging from 2--512.
%
In all configurations, we see that Algorithm \ref{alg:times} provides the fastest performance.
%We observe the same trend found in Figure \ref{fig:time-split-times},

Another metric to use in examining our algorithms is total computational slack time.
%
During MERT, computational slack time arises when some parts of the development set finish translating faster than others.
%
Figure \ref{fig:minmaxspeed} lists the decoding times of the fastest and slowest translation jobs for parts split using each of the four algorithms for various values of $n$, ranging from 2--512.
%
We see the total cumulative slack time for each of these conditions in Figure \ref{fig:slack}.

%Using Algorithm \ref{alg:naive}, the runtimes of the slowest of $n$ translation jobs in each configuration are illustrated in Figure \ref{fig:maxspeed} for various values of $n$ (under the label \textsc{Naive-Split}). Figure \ref{fig:minmaxspeed} shows that in all cases, there is a significant difference between the fastest and slowest translation job, leading to computational slack time (Figure \ref{fig:slack}).

%
%
%
%In nearly all translation models, each sentence is translated independently 
%
%\todo[inline]{Describe Algorithm \ref{alg:words}. \\ \ \\ \ \\ \ \\ \ \\ \ \\ \ }
%
%
%Slack time is essentially eliminated in nearly all ($n<256$) conditions. 
%
%\footnote{Where slack time remains, $nabcheat=(256,512)$, a few jobs were each assigned a single long sentence that took longer to translate than the other jobs which were assigned multiple shorter sentences.}

\begin{figure*}[p]
\centering
\subfloat[Decoding times in \textbf{seconds} for decoder configured using a 5-gram language model.]{
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
\begin{tikzpicture}
\begin{axis}[
xmode=log,
ymode=normal,
xtick pos=left,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={160,150,140,130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Seconds),
ymin=0,
%enlarge x limits=0.5,
%scale only axis=false,
%axis equal=false,
%legend style={
%	at={(0.5,-0.15)},
%	anchor=north,legend columns=-1
%},
%ybar=0pt,
width=13cm,
height=7cm,
ybar,
bar width=7pt,
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
%bar shift=3pt
%ybar interval=0.7
]
%
\input{ngram/ngram.naive-Maximum_times}
\input{ngram/ngram.histogram-Maximum_times}
\input{ngram/ngram.words-Maximum_times}
\input{ngram/ngram.time-Maximum_times}
%
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 1: \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 2: \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 3: \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 4: \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%\caption{\fontspec{TeX Gyre Pagella}Decoding times (in {\textbf{seconds}}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a 5-gram language model.}
%\label{fig:maxspeed}
%\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%

%\qquad
%
%%%%%%%%%%%%%%%%%%%%%%% 
% Begin subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure*}[p]
%\centering
\subfloat[Decoding times in \textbf{hours} for decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.]{
\begin{tikzpicture}%[scale=2]
\begin{axis}[
xmode=log,
ymode=normal,
xtick pos=left,
xtick=data,
ytick=data,
log basis x=2,
xticklabels={8,16,32,64,128,256,512},
ytick={130,120,110,100,90,80,70,60,50,40,30,20,10},
tick label style={font=\tiny},
xlabel=Split Size ($n$),
ylabel=Decode Time (Hours),
ymin=0,
%ybar=0pt,
ybar,
bar width=7pt,
%bar shift=0pt,
width=13cm,
height=7cm,
%cycle list name=black white
%colormap name=bluered
%cycle list={fill={blue!80!red!30}, fill=red!66, fill=yellow!33, black}
%cycle list={fill=blue!30!white,fill=red!60!white,fill=brown!90!white,fill=black}
%cycle list={fill=blue!30!white,fill=red!30!white,fill=brown!30!white,fill=grey}
%cycle list={fill=gray!25,fill=gray!50,fill=gray!75,fill=gray!100}
cycle list={fill=gray!15!blue!15,fill=gray!50!orange!50,fill=gray!70!green!70,fill=gray!100}
%cycle list name=color
]
%
\input{hhmm/hhmm.naive-Maximum_times}
\input{hhmm/hhmm.histogram-Maximum_times}
\input{hhmm/hhmm.words-Maximum_times}
\input{hhmm/hhmm.time-Maximum_times}
%
%\legend{Naive,Histogram,Words,Oracle}
\legend{{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 1: \textsc{Naive-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 2: \textsc{Histogram-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 3: \textsc{Words-Split}},{\fontspec{TeX Gyre Pagella}\footnotesize Algorithm 4: \textsc{Times-Split}}}
\end{axis}
\end{tikzpicture}
}
%%%%%%%%%%%%%%%%%%%%%%% 
% End subfloat
%%%%%%%%%%%%%%%%%%%%%%%
%\caption{\fontspec{TeX Gyre Pagella}Decoding times (in \textbf{hours}) for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}) for a phrase-based decoder configured using a syntactic language model (Schwartz et al., 2011) in addition to a 5-gram language model.}
\caption{\fontspec{TeX Gyre Pagella}Decoding times for the slowest translation job in a translation task split into $n$ decoding jobs using various splitting algorithms (\textsc{Naive-Split}, \textsc{Histogram-Split}, \textsc{Words-Split}, and \textsc{Times-Split}).}
\label{fig:maxspeed}
\end{figure*}

@@

We observe that while Algorithm \ref{alg:histogram} fails to improve over Algorithm \ref{alg:naive} for a standard Moses configuration for small values of $n$, for values of $n>8$, and for all values of $n$ using the slow syntactic language model, Algorithm \ref{alg:histogram} represents a clear improvement. Results for Algorithm \ref{alg:words} show further speedups over Algorithms \ref{alg:naive} and \ref{alg:histogram} in most configurations. Algorithm \ref{alg:times} is the clear winner, nearly eliminating slack time in many cases.

%In Figure \ref{fig:minmaxspeed}a, we observe that the maximum runtimes using Algorithm \ref{alg:histogram} fail to improve over the runtimes using Algorithm \ref{alg:naive} for a standard Moses configuration for small values of $n$. But for values of $n>8$, and for all values of $n$ using the slow syntactic language model, Figure \ref{fig:minmaxspeed}b shows that Algorithm \ref{alg:histogram} results in significant decreases in total decoding time over the naive Algorithm \ref{alg:naive}.
%
%Results for Algorithm \ref{alg:words} show substantial speedups over Algorithms \ref{alg:naive} and \ref{alg:histogram} when using the slower decoder configuration. Similar speedups are seen where $n>8$ for the faster configuration.
%
%This technique results in speedups under all conditions (Figure \ref{fig:minmaxspeed}). In nearly all conditions, the speedup is substantial over the baseline. 
%
%Figure \ref{fig:slack} shows that slack time is at or near zero in all cases where $n<256$.

@@

Implementations in Ruby and Java of all four splitting algorithms are provided at \url{http://github.com/dowobeha/balance-corpus}.

@@

\clearpage
\thanksnostar{Opinions, interpretations, conclusions, and recommendations are those of the authors and are not necessarily endorsed by the United States Air Force. Cleared for public release on 19 Jan 2011. Originator reference number RH-11-107802. Case number 88ABW-2011-0185. Thanks to Grant Erdmann and Tim Anderson, and to the anonymous reviewers.}
